A Formal Model of Checked C

### 2023-01-17 sandp23/evaluation.tex

>>>>

\begin{figure*}[t]
\includegraphics[width=1.0\linewidth]{images/program_memory_benchmark.png}
\caption{Memory benchmarking of ~\systemname.}
\label{fig:memory}
\end{figure*}

\begin{table*}[htb]
  \centering
  \begin{tabular}{|p{2.6 cm }|p{1.6 cm}|p{1.9cm}|p{1.7cm}|p{2cm}|p{1.5cm}|p{1.5cm}|p{1.5cm}|}
    \hline
    \textbf{Task} & \textbf{ProFTPD} & \textbf{MicroHTTPD} & \textbf{UFTPD} & \textbf{LibPNG} & \textbf{Tiny Bignum} & \textbf{parsons (wasm)} & \textbf{parsons (tainted) }\\
    \hline
     \textbf{Pointers Annotated} & 6 & 139 & 146 & 248 & 69 & 364 & 378 \\
     \textbf{Lines Sandboxed} & 0 & 650 & 90 & 0 & 30 & 800 & 0  \\
     \textbf{CVEs fixed} & CVE-2010-4221 & CVE-2013-7039 & CVE-2020-14149, CVE-2020-5204 & CVE-2018-144550 & - & - & -  \\
     \textbf{Time To Port} & 1 Day & 2 Days & 2 Days & 3 Days & 4 Hours & 1 Day & 2 Days \\
    \hline
  \end{tabular}
  \caption{Conversion Efforts}
\end{table*}

\subsection{Security Impact}
\label{subsec:securityimpact}
The last column of~\tbl{TODO} shows the list of all spatial safety vulnerabilities in the functions that have tainted types or are isolated in the~\ucregion{}.
As expected,~\emph{all vulnerabilities} are prevented by~\systemname{}.
The symbols~\vulprevented{} and~\vulisolated{} indicate whether the vulnerability was detected by our dynamic instrumentation or isolated in the sandbox, respectively.
This shows that~\systemname{} provides an effective mechanism to prevent spatial safety vulnerabilities.

\aravind{Fix the conversion effort table by the correct symbols}

\subsection{Program Benchmarks}
\myparagraph{Overview}
We evaluate \systemname on the basis of conversion efforts and performance on six programs with the use of a pre-included test suite for each program. We intend to demonstrate \systemname's capability of enforcing spatial safety in all of the scenarios leading to modern memory vulnerabilities by re-introducing corrected memory bugs and then performing \systemname annotations on the relevant buggy code . However, this requires the bugs to be predicted before they are discovered, which is chronologically impossible. Consequently, we also annotate large sections of relevant bug-free code to mimic the developer's intuition on making \systemname annotations without any input on the bug locality. Consequently, each of the conversions for the six programs follows a varied approach.

\myparagraph{Observations}
Our evaluation of \systemname on six programs yields the following results inferred from Figure ~\ref{fig:runtime} and Figure ~\ref{fig:memory}:
\begin{itemize}
  \item Run-time Overhead is proportional to the extent of 
annotated pointers and sandboxed code.
  \item Marshalling can be avoided if Taintedness of a pointer is propagated throughout across all its usages.
  \item Marshalling data might sometimes be relatively cost-effective instead of propagating the taintedness of a pointer throughout the program.
  \item Marshalling can be avoided by using \_TLIB function qualifier that allows tainted pointers to be passed as parameters to corresponding non-tainted pointer arguments of library functions that are assured by the developer to not leak tainted addresses.
  \item Conversion style and extent dictate the converted program's conversion efforts and is modest at a couple of days. 
  \item Annotating unsafe variables with tainted pointers is more cost-effective than sandboxing the entire function, given the function is assured to be control-flow safe.  
\end{itemize}

### 2023-01-17 sandp23/implementation.tex

\subsection{\sourcerewriter} \label{subsec:checkmate}
The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (\numcheckmatelines{3028}).
However, we use a small OCAML program (\numocamllines{680}) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).
\aravind{Arun, please fix the number of lines you changed for clang frontend and ocaml program}

@@

\noindent\emph{\_Mirror}: 
This qualifier permits copying the corresponding function into both ~\cregion and~\ucregion, which permits the handling of certain simple utility functions that might be called from both regions. For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions. 
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_Mirror int append_string(_TPtr<char> buf,
const char* appendStr : itype(_Nt_array_ptr<const char>),
_TPtr<char> buf_start, size_t buf_len) {
/* Qualifier Rules:
1.) No access to global data NOT marked "const"
2.) Callees must be _Tainted or _Mirror
*/
...
}

\end{minted}
Qualifying ~\code{append_string} with~\code{_Mirror} duplicates the function in both regions,
so that calling \code{append_string} permits the avoidance of over-tainting \code{appendStr},
as \code{appendStr} can be unchecked and checked pointers in $\umode$ and $\cmode$ regions, respectively.
"\_Mirror" enforces control-flow and data-flow compile-time semantic rules as shown above to assert region locality and exclusivity amongst between both the regions. \liyi{fix the sentence.}

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions without memory modification allowing developers to use the function freely in ~\cregion.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the _TLIB annotation below.
strncat(dst, t_str, n);
\end{minted}

@@

Passing tainted pointer \code{t_str} to unqualified ~\code{strncat} above is disallowed without having additional sandboxing for \code{t_str}. If a user ascertains that \code{t_str} has the right buffer size for \code{strncat}, she might label \code{strncat} with \code{\_TLIB}, so that \code{t_str} is treated as an unchecked pointer; such annotation relaxes type-checking for all the arguments to its calls. It is worth noting that \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying; therefore, using \code{\_TLIB} requires users' awareness of memory address leaks.

@@

\subsection{Compiler}
\label{subsec:compilerimple}
The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
We implemented the compiler (\numcompilerlines{}) on top of the existing~\checkedc compiler.
\aravind{Arun, please fix the number of changes you made for the compiler}

Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
For instance, for the code in~\lst{lst:final}, our type checker emits an error saying that ``~\emph{buffer~\code{buff} is tainted, but it is allocated in~\cregion}''.
Consequently, the developer has to make the following changes to allocate~\code{buff} in the sandbox:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
- char buff[MAX_MSG_SIZE] _Tainted;
+ _t_Array_ptr<char> buff: count(MAX_MSG_SIZE)
+    = sbx_allocate(MAX_MSG_SIZE);
...
+ sbx_free(buff)
return rc;
\end{minted}

@@

The~\ucregion source files will be compiled with sandbox compiler,~\ie NAMEOFSANDBOXCOMPILER (CITATION).
Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
\cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers. 
\aravind{Arun, please fill up the name of the sandbox compiler and add citation.}

>>>>

\subsection{\sourcerewriter} \label{subsec:checkmate}
The~\sourcerewriter is primarily implemented in C++ as a~\clang frontend tool (3028 SLoc).
However, we use a small OCAML program (680 SLoc) to remove annotations and make the code compilable with the sandbox compiler (\sect{subsubsec:genucregion}).

@@

\noindent\emph{\_Mirror}: 
This qualifier permits copying the corresponding function into both ~\cregion and~\ucregion, which permits the handling of certain simple utility functions that are called from both regions. For example, ~\code{append_string} in our evaluation of parsons\_wasm has callers from both the regions. 
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
_Mirror int append_string(_TPtr<char> buf,
const char* appendStr : itype(_Nt_array_ptr<const char>),
_TPtr<char> buf_start, size_t buf_len) {
/* Qualifier Rules:
1.) No access to global data NOT marked "const"
2.) Callees must be _Tainted or _Mirror
*/
...
}

\end{minted}
Qualifying ~\code{append_string} with~\code{_Mirror} duplicates the function in both regions, allowing calls to \code{append_string} with parameter to \code{appendStr} as an unchecked or checked pointer within $\umode$ and $\cmode$ regions, respectively. Consequently complexity from over-tainting is avoided as ~\code {appendStr} need not be tainted in ~\cregion and neither are callbacks required to access ~\code{append_string} from ~\ucregion.   
"\_Mirror" enforces control-flow and data-flow compile-time semantic rules to ensure all variable and function call dependencies of mirrored functions required for ~\ucregion's compilation are resolved.

\noindent\emph{\_TLIB}:
This qualifier relaxes type-checking rules on library functions, allowing developers to use the function freely in ~\cregion.
\begin{minted}[mathescape, escapeinside=||, fontsize=\footnotesize]{c}
// First, manually check the memory is in tainted region.
// if yes, then call strncpy.
if (!is_mem_in_range(t_str, t_str + n, SBX_LOW(), SBX_HIGH()))
  handle_violation();
// our type checker ignores this because 
// the _TLIB annotation below.
strncat(dst, t_str, n);
\end{minted}

@@

Passing tainted pointer \code{t_str} to unqualified ~\code{strncat} above is disallowed without having additional ~\ucregion implementation for \code{strncat}. If a user ascertains that \code{t_str} has the right buffer size for \code{strncat}, she might label \code{strncat} with \code{\_TLIB}, so that \code{t_str} can be treated as an checked pointer parameter; such annotation relaxes type-checking for all the arguments to its calls. It is worth noting that \systemname does not enforce any semantics to ensure ~\code{\_TLIB} functions implemented in ~\cregion are non memory-modifying; therefore, using \code{\_TLIB} requires users' awareness of memory address leaks.

@@

\subsection{Compiler}
\label{subsec:compilerimple}
The~\systemname compiler includes the type checker (\sect{sec:typechecking}) and necessary instrumentation to handle tainted pointers (\ie from sandbox) and perform runtime checks.
We implemented the compiler as a clang library (1812K SLoc) on top of the existing~\checkedc compiler.
Our type checker verifies the constraints (\fig{fig:type-system-1}) and displays violations as errors, enabling developers to make additional annotations.
For instance, for the code in~\lst{lst:final}, our type checker emits an error saying that ``~\emph{buffer~\code{buff} is tainted, but it is allocated in~\cregion}''.
Consequently, the developer has to make the following changes to allocate~\code{buff} in the sandbox:
\begin{minted}[mathescape, escapeinside=||, fontsize=\tiny]{diff}
- char buff[MAX_MSG_SIZE] _Tainted;
+ _T_Array_ptr<char> buff: count(MAX_MSG_SIZE)
+    = sbx_allocate(MAX_MSG_SIZE);
...
+ sbx_free(buff)
return rc;
\end{minted}

@@

The~\ucregion source files will be compiled with sandbox compiler,~\ie WASI-SDK (~\cite{wasi-clang}).
Given that WASM sandbox uses 32-bit addresses (v/s 64-bit for~
\cregion), our instrumentation also adds necessary pointer swizzling to have transparent access to composite data types,~\ie~\code{struct}'s with pointers.
